\Title

### 2020-12-10 manuscript/manuscript.tex
Write gradient boosting as an algorithm

\documentclass[twocolumn]{article}
\usepackage[left=0.7in,right=0.7in,top=1in,bottom=1in]{geometry}
\setlength{\columnsep}{2\columnsep}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[round,authoryear,sort]{natbib}
\usepackage{url}
\usepackage[pdftex,colorlinks=true]{hyperref}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{booktabs}

@@

Here we propose an iterative algorithm for fitting the source coefficients
using a gradient boosting \citep{friedman2001, friedman2002} strategy:
(i) define a collection of overlapping windows of equal size that cover the
survey area,
(ii) define one set of equivalent sources per window,
(iii) iteratively fit the coefficients of each set of sources.
Once every set of sources is fitted, we can predict the value of the field on
any observation point as the effect of every set of equivalent sources.

Lets assume we have $N$ measurements of a harmonic field on observations points
$\{\mathbf{p}_i\}_{i=1}^{N}$ stored in the vector $\mathbf{d} = (d_1, d_2,
\ldots, d_N)^T$. We have defined a set of $M$ point sources located on
$\{\mathbf{q}_j\}_{j=1}^{M}$. Instead of trying to fit the entire set of
coefficients $\mathbf{c}$ at once, we can reduce the problem to fitting $K$
different sets of equivalent sources: we define $K$ overlapping windows of
equal size that cover the whole observation area, and then we create one set of
equivalent sources per window formed by the $\{\mathbf{q}_m\}_{m=1}^{M_k}$
sources that fall inside the $k$-th window.
The observed data should be equal to the superposition of the
effect of every set of equivalent sources:

\begin{equation}
    \mathbf{d} = \sum_{k=1}^K \alpha_k \mathbf{A}_k \mathbf{c}_k,
\end{equation}

\noindent where $\mathbf{c}_k$ is the vector of coefficients for the $k$-th
set of equivalent sources; $\mathbf{A}_k$ is the Jacobian matrix between
those sources and every data point; and each $\alpha_k$ is a constant called
\emph{step-size parameter}.

@@

\begin{equation}
    \alpha_k = \frac{
        \mathbf{r}_{k-1}^T \mathbf{A}_k \mathbf{c}_k
    }{
        \left(\mathbf{A}_k \mathbf{c}_k\right)^T \mathbf{A}_k \mathbf{c}_k
    }.
\end{equation}

>>>>

\documentclass[twocolumn]{article}
\usepackage[left=0.7in,right=0.7in,top=1in,bottom=1in]{geometry}
\setlength{\columnsep}{2\columnsep}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[longend,linesnumbered]{algorithm2e}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[round,authoryear,sort]{natbib}
\usepackage{url}
\usepackage[pdftex,colorlinks=true]{hyperref}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{setspace}

@@

Gradient boosting was first introduced by \citet{friedman2001, friedman2002} as
a method for fitting additive parametric models of the form

\begin{equation}
  d = \sum_{k=1}^K \alpha_k f(\mathbf{c}_k),
\end{equation}

\noindent where $\alpha_k$ is a scalar coefficient called the \emph{step-size}
and $f$ is a function of the parameter vector $\mathbf{c}_k$.
For linear regression problems, these additive models can be written as the
matrix equation

\begin{equation}
    \mathbf{d} = \sum_{k=1}^K \alpha_k \mathbf{A}_k \mathbf{c}_k \ .
    \label{eq:gb-linear-model}
\end{equation}

We can transform our equivalent source problem in
equation~\ref{eq:linear-problem} into an additive model by
(1) defining a set of $K$ overlapping windows of equal size that cover the
survey area and (2) defining separate sets of equivalent sources, one per
window. Then, the coefficients $\mathbf{c}_k$ relate to the equivalent sources
of the $k$-th window and $\mathbf{A}_k$ is the Jacobian matrix between
those sources and every data point.
Thus, the predicted data are a superposition of the
effects of the $K$ sets of equivalent sources.

The gradient boosting algorithm for finding the damped least-squares solutions
for the $K$ parameters vectors $\mathbf{c}_k$ in
equation~\ref{eq:gb-linear-model} can be modified from \citep{friedman2001}

\begin{algorithm}[!h]
  \DontPrintSemicolon
  \setstretch{1.5}
  define the residual vector: $\mathbf{r}_{0} = \mathbf{d}^o$ \;
  \For{ $k = 1$ \KwTo $K$ }{
    $\mathbf{B}_k = \mathbf{A}_k \mathbf{S}_k\inv$ \;
    $
     \hat{\mathbf{m}}_k = \left[\mathbf{B}_k\trans \mathbf{W}_k \mathbf{B}_k +
     \lambda \mathbf{I} \right]\inv \mathbf{B}_k\trans \mathbf{W}_k
     \mathbf{r}_{k-1}
    $ \;
    $\hat{\mathbf{c}}_k = \mathbf{S}_k\inv \hat{\mathbf{m}}_k$ \;
    $
      \alpha_k = \underset{\alpha}{\mathrm{argmin}}
      \left(
        \left\lVert
          \mathbf{r}_{k - 1} - \alpha \mathbf{A}_k \hat{\mathbf{c}}_k
        \right\rVert ^ 2
      \right)
    $ \;
    $\mathbf{r}_k = \mathbf{r}_{k - 1} - \alpha_k\mathbf{A}_k \hat{\mathbf{c}}_k$ \;
  }
  \BlankLine
  \setstretch{1}
  \caption{Gradient boosting solution for damped least-squares regression.}
\end{algorithm}

The algorithm above works by fitting each component of the additive model, one
at a time, to the residuals of the previous component.

@@

\begin{equation}
    \alpha_k = \frac{
      \mathbf{r}_{k-1}\trans \mathbf{A}_k \hat{\mathbf{c}}_k
    }{
      \hat{\mathbf{c}}_k\trans\mathbf{A}_k\trans \mathbf{A}_k \hat{\mathbf{c}}_k
    }.
\end{equation}
